Tables and Figures

### 2022-02-28 analysis/paper/abstract.Rmd
add S prefix to all tables and figs in supp mat

@@

@@

<!-- Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. Under these circumstances, the length of the paragraph can be yp to 300 words. (This example is 190 words without the final section, and 250 words with it).

>>>>

---
output:
  bookdown::pdf_document2:
    keep_tex: true
---

@@

```{r include=FALSE}
library(here)
library(patchwork)
source(here("analysis/scripts/analysis.R"))
source(here("analysis/scripts/tables.R"))
perc_tbl <- prop %>%
  ungroup() %>%
  select(!c(sample, solution, prop)) %>%
  pivot_wider(names_from = size, values_from = perc) %>%
  add_row(perc_comb_row, .after = 2)
# prepare cor output for reporting
cor_r <- signif(count_cor$estimate, 3)
cor_ci <- signif(count_cor$conf.int, 3)
cor2_r <- signif(starch_cor$estimate, 3)
cor2_ci <- signif(starch_cor$conf.int, 3)

#w_count <- wordcountaddin::word_count()
```

@@

<!-- Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. -->

Dental calculus has proven to contain a wealth of information 
on the dietary habits of past populations. These insights have, to a large extent,
been obtained by the extraction and identification of
starch granules contained within the mineralised dental plaque from a wide range 
of regions and time periods. The scope of previous studies have been limited to 
microfossil extraction and identification to reconstruct dietary preferences from
the archaeological record, and few studies have attempted to address the biases 
of starch retention in dental calculus. Those that have considered this problem 
have been limited to *in vivo* studies on modern humans and non-human primates.
Here, we present a multispecies oral biofilm model, which allows experimental 
research on  starch incorporation and retention to be conducted on *in vitro* 
dental calculus in a controlled laboratory setting. The biofilms were exposed 
to treatment solutions with known quantities of
dietary starches (wheat and potato) during the 25-day growth period. After this,
the starch granules were extracted from the mature biofilm (by dissolution in EDTA),
and counted. We show that the granule counts extracted from the model dental calculus 
represented a low proportion (ranging from `r paste(min(perc_tbl$total), max(perc_tbl$total), sep = " to ")`) of the total number 
of granules exposed to the biofilms throughout the experiment.
Additionally, we found that the ratios of granule sizes from the extracted 
starch granules differed from the original treatment solutions, with large 
granules (>20 $\mu$m) consistently being under-represented.
We also found a correlation
between the absolute granule counts and dry-weight of the biofilm 
(*r* = `r cor_r`, 90%CI[`r paste(cor_ci[1], cor_ci[2], sep = ", ")`]), as well
as between the concentration (count per mg) of granules and dry-weight 
(*r* = `r cor2_r`, 90%CI[`r paste(cor2_ci[1], cor2_ci[2], sep = ", ")`]).  
Our results reinforce previous *in vivo* studies suggesting that dental calculus 
presents a very small, and partly biased picture of the original dietary intake 
of starches, with an over-representation of plants producing granules smaller than
20 $\mu$m in size. The experimental model presented here is well-suited to address
the need for further validation of methods and biases associated with dietary
research on dental calculus.

### 2022-02-28 analysis/paper/fig_tab.Rmd
add S prefix to all tables and figs in supp mat

## Tables
<!-- ### Tables ### -->

@@

```{r sample-count}
count_sample <- count_sample %>%
  ungroup() %>%
  add_row(count_sample_comb, .after = 2)

# convert columns to contain mean count and percentage of total count
count_samp_per <- count_sample %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(signif(x / total, 3)), ")")))

count_samp_out <- tibble(count_samp_per, count_sample_sd, .name_repair = "unique") %>%
  select(1,2,3,7,4,8,5,9,6,10) #%>%

count_samp_out[1,3] <- NA # replace with NA
  #add_row(tibble_row(count_sample_comb_out), .before = 3)
```

```{r solution-count-out}
count_sol_out <- count_sol %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

### adding the 'both' row to the data frame ###
count_sol_comb_out <- count_sol_comb %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

count_sol_out <- count_sol %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")"))) %>%
  ungroup() %>%
  add_row(tibble_row(count_sol_comb_out), .before = 3)

count_sol_out[1,3] <- NA
```

```{r solution-count-tbl}
cnames <- c("Solution", "Starch", "Small (%)", "Medium (%)", "Large (%)", "Total (%)")
options(knitr.kable.NA = "")
knitr::kable(count_sol_out, col.names = cnames, 
             caption = "Mean starch counts from solutions, including the
             proportional makeup of the different sizes of granules.")
```

```{r sample-count-tbl}
cnames <- c("Treatment", "Starch", "Small (%)", "SD", "Medium (%)", "SD", 
            "Large (%)", "SD", "Total (%)", "SD")
knitr::kable(count_samp_out, col.names = cnames, 
             caption = "Mean starch counts extracted from samples with standard 
             deviation (SD), including the proportion of granule sizes of the total count.")
```

```{r sample-perc}
# convert proportions to percent labels
perc_comb <- prop_comb %>% 
  mutate(across(c(s, m, l, total), scales::percent, 0.001))

perc <- prop %>%
  mutate(across(c(s, m, l, total), scales::percent, 0.001)) %>% 
  add_row(tibble_row(perc_comb), .before = 3)
```

```{r sample-prop-tbl}
cnames <- c("Treatment", "Starch", "Small", "Medium", "Large", "Total")
knitr::kable(perc, col.names = cnames, 
             caption = "The mean percentage of starches from the solutions that 
              were incorporated into the samples.")
```

## Figures
<!-- ### Figures ### -->

@@

```{r cor-plot, warning=FALSE, fig.cap="Scatter plot of sample weight and standardised starch count by Z-score for separated treatments."}
pl_cor
#pl_cor + pl_cor2 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```

```{r cor-plot2, warning=FALSE, fig.cap="Scatter plot of sample weight in mg and standardised count of starch grains per mg calculus."}
pl_cor2
```

>>>>

# Tables

@@

```{r solution-count-tbl}
cnames <- c("Solution", "Starch", "Small (%)", "Medium (%)", "Large (%)", "Total (%)")
options(knitr.kable.NA = "")
knitr::kable(sol_tbl, col.names = cnames, 
             caption = "Mean starch counts from solutions, including the
             proportional makeup of the different sizes of granules.")
```

```{r sample-count-tbl}
cnames <- c("Treatment", "Starch", "Small (%)", "SD", "Medium (%)", "SD", 
            "Large (%)", "SD", "Total (%)", "SD")
knitr::kable(samp_counts_tbl, col.names = cnames, 
             caption = "Mean starch counts extracted from samples with standard 
             deviation (SD), including the proportion of granule sizes of the total count.")
```

```{r sample-prop-tbl}
cnames <- c("Treatment", "Starch", "Small", "Medium", "Large", "Total")
knitr::kable(perc_tbl, col.names = cnames, 
             caption = "The mean percentage of starches from the solutions that 
              were incorporated into the samples.")
```

# Figure Captions

@@

```{r cor-plot, warning=FALSE, fig.cap="Scatter plots of (A) sample weight in mg and standardised starch count by z-score for separated treatments, and (B) sample weight in mg and standardised count of starch grains per mg calculus."}
#pl_cor
pl_cor + pl_cor2 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
#ggsave("frontiers/figures/cor-plot.png", dpi = 300)
```
